### 2020-12-04 article_rvTH15_6_essai.tex

\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage{setspace}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\doublespacing
\section*{Question Part 1.}
I am writing this document inside RStudio, using Sweave. I want everything 
in the document to be doublespaced with the exception of R code chunks, 
which I want to be singlespaced for the sake of tidiness and saving space. 
It is important that the code chunks also run in R, since they may be doing 
things like telling R to draw plots which will be part of the document.
\section*{Question Part 2.}
The problem that I seem to be having is that when I put in a command to 
begin singlespacing and a command to end single spacing around an R code 
chunk, text that is not inside the begin and end singlespacing command, for 
some reason, ends up being singlespaced as well, even though as far as I can 
tell it should be doublespaced, as can be seen with the text in this very 
paragraph (if you've compiled it into a .pdf).

\begin{minipage}{\linewidth}
\singlespacing
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{k}\hlkwb{=}\hlnum{3}
\hlstd{mse}\hlkwb{=}\hlnum{0.347}
\hlstd{d}\hlkwb{=}\hlnum{0.85}
\hlstd{alpha}\hlkwb{=}\hlnum{0.05}
\hlstd{n}\hlkwb{=}\hlnum{30}
\hlstd{pwr.df}\hlkwb{=}\hlkwd{data.frame}\hlstd{(}\hlkwc{row.names} \hlstd{=} \hlnum{1}\hlopt{:}\hlstd{n)}
\hlstd{pwr.df}\hlopt{$}\hlstd{n}\hlkwb{=}\hlnum{1}\hlopt{:}\hlstd{n}
\end{alltt}
\end{kframe}
\end{knitrout}
\end{minipage}
Strangely, the problem doesn't occur for text which follows the "end" of the 
singlespacing commmand, but text which comes before the 'begin' 
singlespacing command is affected, as can be seen here.
\end{document}