### 2020-12-13 article_rvTH17.Rnw
version17

\documentclass[article]{jss}

%% -- LaTeX packages and custom commands ---------------------------------------

%% recommended packages
\usepackage{thumbpdf,lmodern}
\usepackage{float,tikz}
%% another package (only for this demo article)
\usepackage{framed}
%% author pakckage
\usepackage{xcolor,caption, float}
\usepackage{array}

\graphicspath{ {./imgs/} }

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}

%% This modifies which proportion of the page must be occupied by the figure
%% or table to take the full page (with no additional text).
%% Check if the journal does not have any rule about not changing this.
\renewcommand{\floatpagefraction}{.8}

% \setlist[itemize]{noitemsep, topsep=0pt}
<<preliminaries, echo=FALSE, include=FALSE, cache=FALSE >>=
# options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
library("MASS")
library("extrafont") #;library("ggplot")
library("igraph")
opts_chunk$set(prompt=TRUE, fig.pos = '!htbp', results='asis', highlight=FALSE,
               class.source = "bg-success")
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
options(SweaveHooks=list(fig=function() par(mar=c(0, 0, 0, 0))))
# options(replace.assign=TRUE, width=90, prompt="R> ")
opts_knit$set(concordance=TRUE, self.contained=FALSE, background = "#FFFFFF")
# knitr::opts_chunk$set(fig.pos = '!htbp')
loadfonts(device = "win")
@

%% -- Article metainformation (author, title, ...) -----------------------------

%% - \author{} with primary affiliation
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Thomas Huet\\UMR 5140 \And Jose Pozo \And Craig Alexander}
%   \And Second Author\\Plus Affiliation}
\Plainauthor{Thomas Huet, Jose Pozo, Craig Alexander}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{Modelling Prehistoric Iconographic Compositions. The \proglang{R} package \pkg{decorr}}
\Plaintitle{Modelling Prehistoric Iconographic Compositions. The R package 'decorr'}
\Shorttitle{Modelling Prehistoric Iconographic Compositions}

%% - \Abstract{} almost as usual
\Abstract{
By definition, Prehistorical societies are characterised by the absence of a writing system. During, the largest part of human history -- from far -- symbolic expressions belong to illiterate societies which express themself with rock-art paintings, pottery decorations, figurines and statuary, etc., and a lot of now disappeared carved woods, textil design, etc. At the composition level, recognition of meaningful associations of signs and reccurent patterns indicate clearly the existence of social conventions in the way to display and to read these expressions.
We present the \pkg{decorr} \proglang{R} package which grounds concepts, methods and tools to analyse any ancient graphical systems. Our asumption is that i) any graphical system is a spatial distribution of features, and ii) these features have possibly any meaningful relationships one with another depending on their pairwise spatial proximities. To model the graphical compositions we employ concepts coming from the Graph Theory. To ensure the feasability of this type of analysis, we propose a GIS-based method for inputs and a serie of functions for data management.
}

\Keywords{Semiotics, Iconography, Prehistory, Graphs, Spatial Analysis, Binary Topological Relationships (\emph{birel}), \proglang{R}}
\Plainkeywords{Semiotics, Iconography, Prehistory, Graphs, Spatial Analysis, Binary Topological Relationships (birel), R}

\Address{
Thomas Huet\\
CNRS-UMR 5140 \\
Archeologie des Societes Mediterraneennes\\
Universite Paul Valery\\
route de Mende\\
Montpellier 34199, France\\
E-mail: \email{thomashuet7@gmail.com}\\
%  URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
}

\begin{document}
%\SweaveOpts{concordance=TRUE}
% \SweaveOpts{concordance=TRUE}
% \Sweave2knitr("article_rvTH14_1.Rnw")
% \SweaveOpts{concordance=TRUE}

\section[Introduction]{Introduction} \label{sec:intro}

Symbolic practices is a characteristic trait of human societies. Even discussed, such practices seems to start between 233,000 to 800,00 BC \citep{dErrico00}, covering more than 97\% of total human societies time span. Symbolism cover a large range of practices, from ochre deposit in a tomb, to menhir alignements, passing through wall fresco. This latter, what might be called "iconographical practices", probably shows the most complex and interesting testimonies of past societies symbolism. For decades, its study was linked to history of religion because commonly seen as closely linked to cultual practices and believes. Since the \textit{New Archaeology} developpement during the 60's \citep{Clarke14}, symbolic expressions start to be studied with the same formal methods (statistics, seriations, distribution maps, etc.) as any another aspect of social organisation: settlement patterns, tools \emph{chaine op\'eratoire}, susbsitence strategies, etc. \citep{Renfrew91}. But unlike many aspects of the material culture where technological requirements and technical efficiency determine the choice of the raw material and the shape of the object -- a flint blade for cutting, a pottery for containing, a house for living --, the function of an iconographic composition cannot be drawn directly from itself. Furthermore, there is no specific and controled vocabularies to describe hand-made images. There is always a difficulty to decribe an iconographical content with alphanumerical signs. Whether these last decades study of ancient iconography had  undergone significative improvements at the site scale -- with GIS/database statistics, paleoenvironmental restitutions, etc. -- and at the sign scale with the development of archaeological sciences -- radiocarbon dating, use-wear analysis, elemental analysis, etc. --, these improvement do not necessarly help to understand the semantic content of the iconography.\\
Semantics or semiotics can be defined as a system of conventional features -- called signs -- organised also in conventional manners. Until our days, formal methods to study ancient iconography semantics have been mostlty been grounded  -- explicitly or not -- on the prime principle of Saussurian linguistic: the '\emph{linearity of the signifier}' \citep{Saussure89}.
Writing is one of the most rational semiographical system with a clear distinction between signified and signifier and the development of the signified on a linear axis. Even if we do not understand the meaning of the signs, writing can easily be modelled with Graph theory and recurent patterns can be identified. For example, the 3-letters word "\code{art}" can be modelled with three vertices (\code{a}, \code{r}, \code{t}) and two edges (one between \code{a} and \code{r}, the other between \code{r} and \code{t}). In \proglang{R}, these features, concatenated in this order with a \code{paste0()}, is \code{art}, and not \code{rat} (Fig. \ref{fig:fart}).

<<a_r_t, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=1.25, fig.height=0.25, fig.cap="\\label{fig:fart}Concatenate of 'a', 'r' and 't' graphical units is 'art'">>=
par(mar=c(0.1,0.1,0.1,0.1) )
g <- graph_from_literal(a-+r, r-+t)
plot(g,
     layout = cbind(seq(-1,1,length.out = gorder(g)), 0),
     vertex.color="white",
     vertex.frame.color="white",
     vertex.size=40,
     vertex.label.cex=1,
     vertex.label.color="black",
     vertex.label.family="Courier",
     edge.arrow.size=0.5,
     xlim=c(-0.94,0.94), 
     ylim=c(-1,-1))
@

But, as stated, in Prehistorical the writing system does not exists. Spatial relationships between graphical features, or graphical units, are not necessarly linear and directed but could most probably be more multi-directional and undirected: the direction of the interactions of pairwise graphical units can be in any order (Fig. \ref{fig:fratar}).

<<echo=FALSE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=1.4, fig.height=1.4, fig.cap="\\label{fig:fratar}Potential spatial relations between graphical units">>=
par(mar=c(0,0,0,0))
ugs <- data.frame(name=c("a","r","t"))
lks <- data.frame(from=c("a", "r", "t"), to=c("r", "t", "a"), type=c("=","=","="))
g <- igraph::graph_from_data_frame(lks, directed=FALSE, vertices= ugs)
plot(g,
     vertex.color="white",
     vertex.frame.color="white",
     vertex.size=80,
     vertex.label.cex=1,
     vertex.label.color="black",
     vertex.label.family="Courier",
     xlim=c(-0.65,0.65), 
     ylim=c(-0.65,0.65)
)
@

Because of the inherent variability of iconography, and because graphical and spatial proximities between graphical units are generaly not quantified, applying the Saussurian model to any prehistorical graphical content had led to considerable problems:
\begin{itemize}
\setlength\itemsep{.05em}
\item unexplicit spatial groupings of graphical units -- like graphical units grouped into \emph{figures},  \emph{figures} grouped into \emph{patterns}, \emph{patterns} grouped into \emph{motives}, etc. -- with tedious number of groups
\item consistency, proximities and relationships between these groups are often implicit and not quantified
\item studies develop proper descriptive vocabularies, singular relationships of grouping, idosyncratic methods at site-dependent or period-dependent scales
\end{itemize}

These issues limit drastically the possibility to conduct cross-cultural comparisons and to draw a synthesis of humankind's symbolism at a large scale and over the long-term.

According to us, prehistorical composition -- like any formal system -- are spatial features (mostly shapes) related one with the other depending on rules of spatial proximities. Then, the First Law of Geography: "\emph{everything is related to everything else, but near things are more related than distant things}" \citep{Tobler70} should be more operative than the Saussurian '\emph{linearity of the signifier}' principle \citep{Saussure89}. To get these '\emph{near things}', or features proximities, we use concepts coming from the Graph Theory, which offers a general framework to manage neighbourhoods. Furthermore, we address the questions of managing graphical units attributes (for example an anthropomorph with a sword \emph{vs} an anthropomorph with a spear) and superimpositions of graphical units.  
% and particularly relative neighborhood graphs (RNG) which is '\emph{a powerful model of low-level visual process}' \citep{Toussaint80}

In this article we present the \proglang{R} package \pkg{decorr}. Its purpose is to formalise a method and offer tools to analyse any graphical content (here: decorations) with methods coming from raph Theory, GIS and binary topological relationships (\emph{birel}) modeling. This package has been grounded on the seminal work of \citet{Alexander08} and its first IT implementation by \citet{Huet18a}.

\section[Model]{Graph theory Model} \label{sec:model}

Graph theory offers a conceptual framework and indices (global at the entire graph scale, local at the vertex scale) to deal with notions of networks, relationships and neighbourhoods. Graphical units (GUs) can be modelled as vertices (nodes) separated one with another by undecorated areas. Their spatial relations can be modelled as edges. The different levels of GUs spatial organisation can be retrieve by graphs analysis (Graph Theory) and a spatial (GIS) analysis.

Graph elements, nodes and edges -- repectively GUs and connexions between GUs -- are created on a GIS interface.  
Indeed, for large series of decorated supports, GIS offers the most suitable and flexible interface to register all GUs and to get their coordinates, makes easy the updating of GUs and GUs relation and opens to develop further spatial analysis.  The decoration image is considerated as the basemap of the project and will cover the region of interest of the analysis (ie, all the graphical content). The decoration image can be binarized: GUs are considerated active, the undecorated parts of the support -- the background -- are considerated inactive. The polygonisation of the GUs (\code{POLYGONS}), including the border of the stelae, allows to calculated their Voronoi seeds (ie, their centroid) and Voronoi cells (ie, Thiessen polygons, their area of influence). So, the entire decoration image is tiled and condidered as a Voronoi diagram. Centroids of the GUs (\code{POINTS}) are calculated and considered as the graph nodes, their locations (\emph{x} and \emph{y} coordinates, measured in pixels) are relative to each decoration. Exist edges (\code{LINES}) between pairs of GUs when their Voronoi cells share a common border (\emph{birel}: 'touches').

The entire graphical content of the decoration is mapped with a geometric graph -- or planar graph -- of GUs centroids (nodes) and GUs proximity links (edges). So, a decoration graph is a one component graph. This graph is the dual graph of the decoration Voronoi diagram (Fig. \ref{fig:gis}).

\begin{figure}[H]
% \begin{figure}[!htbp]
\centering
\includegraphics{gis}
\caption{\label{fig:gis} GIS interface. A) Original decoration of the Late Bronze Age Cerro Muriano 1 stelae (drawing: \cite{DiazGuardamino10}) opened in new GIS project without any projection system with its extent (\code{xmin, xmax, ymin, ymax} in pixels); B) Polygonisation of the GUs, calculation of their centroids (nodes), calculation of the Voronoi cells, calculation of the proximity links (edges); C) plot of the relative \emph{x} and \emph{y} coordinates of the GUs centroids; D) plot of the values in the \code{type} of the GUs}
\end{figure}

Whether this theoritical process include image binarization, polygonization and \emph{birel} analysis, a simplier solution will be to create directly centroids on the GUs and to draw the edges manually. Then \emph{x} and \emph{y} coordinates of nodes can be easily retrieved.\\

\subsection[elements]{Graph elements} \label{sec:elements}

\smallbreak
\textbf{Nodes types}
\\
At least one variable, like the \code{type} of the nodes, is defined in order to analyse and compare iconographical compositions. In this article, we will only consider nodes \code{type} values ('anthropomorph', 'sword', etc.) but it the package user can add other variables (\code{technique}, \code{color}, etc.)

\smallbreak
\textbf{Egdes types}
\\
The package considers the most common types of relations between nodes:

\begin{description}
\item[\emph{normal edges}]
Two contemporaneous nodes having a Voronoi cell sharing a border have a common edge. By convention, this edge is tagged \code{'='} (\code{type = '='}), represented with an orange plain line and its textual notation is \code{'-=-'}. For example: \code{personnage -=- bouclier} means that the nodes \code{personnage} and \code{bouclier} have a common border (Fig. \ref{fig:gis_1}). As said, in Prehistorical iconography, the read direction is not known, so this edge is undirected, ie \code{personnage -=- bouclier} is equal to \code{bouclier -=- personnage} (Fig. \ref{fig:gis_1}).

\item[\emph{attribute edges}]
It occurs frequently that a node can be divided into a main unit (eg, an anthropomorph) and one or various attribute units (eg, a helmet, a male sex). To record this ordered set, the edge between a main and an attribute node is tagged \code{'+'} (\code{type = '+'}), represented with an orange dashed line and its textual notation is \code{'-+-'}. For example \code{personnage -+- casque} means that the main node \code{personnage} has the attribute node \code{casque}. The edge is directed since \code{personnage -+- casque} (the anthropomorph has an helmet) is not equal to \code{casque -+- personnage} (the helmet has an anthropomorph) (Fig. \ref{fig:gis_1}).

% \begin{figure}[H]
\begin{figure}[!htbp]
\centering\emph{}
\includegraphics[height=7cm, trim=15mm 20mm 15mm 20mm, clip, keepaspectratio]{gis_1}
% \includegraphics[height=10cm, trim=15mm 20mm 15mm 20mm, clip, keepaspectratio]{gis_1}
\caption{\label{fig:gis_1} GIS interface, Cerro Muriano 1 decoration. The nodes \code{casque} (helmet) and \code{sexe\_masculin} (male sex) are two nodes attributes of the node \code{personnage} (anthropomorph).}
\end{figure}

\item[\emph{overlapping edges}]
Finally, it occurs that a graphical composition shows decoration phases. These decoration phases can be identified by observing the superimpositions between different nodes or by other means (differences in the techniques or in the patina for example). In the first case, two nodes belonging to two different phase will have a common part (the overlapped intersection) while in the second case the two nodes will not necessarly share a part of their surface. To record this relative chronology, an edge between two non-contemporaneous nodes is tagged \code{'>'} (\code{type = '>'}), represented with a blue plain line and its textual notation is \code{'->-'}. For example \code{ecriture ->- bouclier} means that \code{ecriture} is more recent than the \code{bouclier}, or \emph{overlaps} the\code{bouclier} in the stratigraphical sense. The edge is directed since \code{ecriture ->- bouclier} (ie, \code{ecriture} \emph{overlaps} \code{bouclier}) is not equal to \code{bouclier ->- ecriture} (ie, \code{bouclier} \emph{overlaps} \code{ecriture}) (Fig. \ref{fig:gis_2}).

\end{description}

\begin{figure}[H]
% \begin{figure}[!htbp]
\centering
\includegraphics[height=8cm, trim=30mm 16mm 30mm 13mm, clip, keepaspectratio]{gis_2}
\caption{\label{fig:gis_2} GIS interface, Ibahernando decoration. The node \code{ecriture} (writing) overlaps \code{lance} (spear) and the upper part of the \code{bouclier} (shield).}
\end{figure}

This stratigraphical information (node 1 \emph{over} node 2, or node 2 \emph{under} node 1) helps to understand the relative chronology between node and must be recorded. The analysis of the iconographical content can be stratified (stratified analysis) and performed on each different phases of decoration separately.

The nodes and edges equalities are resumed in the Table \ref{tab:ex.comm}.

<<equalities, results='asis', echo=FALSE, fig.pos = '!htbp'>>=
caption <- paste0("Example of equalities and differences for nodes and edges ")
df.ex <- data.frame(type = c("node", "node", "edge", 
                             "edge", "edge", "edge", "edge"),
                    part.1 = c("epee", "epee", "epee --=-- lance",
                               "epee --=-- lance", "personnage --=-- bouclier", 
                               "personnage --+-- bouclier", "ecriture -->-- bouclier"),
                    part.2 = c("epee", "lance", "epee --=-- lance",
                               "lance --=-- epee", "personnage --+-- bouclier",
                               "bouclier --+-- personnage", "bouclier -->-- ecriture"),
                    is.equal = c("True", "False", "True", 
                                 "False", "False", "False", "False"),
                    stringsAsFactors = F)
print(xtable::xtable(df.ex,
                     caption = caption,
                     label = "tab:ex.comm",
                     size = 7),
      include.rownames = FALSE,
      table.placement = "!htbp")
@

These different relationship types are resumed with their \emph{birel} equivalencies and chronological assumptions in Table \ref{table:tedges}.

% \begin{table}[!htbp]
\begin{table}[H]
\centering
\begin{tabular}{ccccp{2cm}p{4.9cm}}
\hline
% node A & type of edge & node B & undirected & birel & stratigraphical meaning \\
Node 1 & Edge type & Node 2 & Edge directness & \emph{Birel} & Description \rule{0pt}{1em} \\[1pt]
\hline \rule{0pt}{1em}%
\code{A} & \code{=} & \code{B} & Undirected & $ A \cap B = \emptyset $ & 
\code{A} and \code{B} are contemporaneous and disjoint. \\
\code{A} & \code{+} & \code{B} & Directed & $ A \cap B = A $ & 
\code{A} and \code{B} are contemporaneous. \code{B} is an attribute of \code{A}. \\
\code{A} & \code{>} & \code{B} & Directed & $ A \cap B = \exists $ \newline or \newline $ A \cap B = \emptyset $ & 
\code{A} is more recent than \code{B}. \\
% \code{1} & = & 2 & undirected & A \cap B = \varnothing & 1 and 2 can be contemporaneous \\
\hline
\end{tabular}
\caption{Synthesis for the different types of relations between GUs.}\label{tab1}
\label{table:tedges}
\end{table}

\section{The R package decorr} \label{sec:models}

<<results='asis', include=FALSE, echo=FALSE, warning=FALSE>>=
library("decorr")
library("igraph")
library("magick")
@
%
The \pkg{decorr} package can be downloaded from GitHub
%
\begin{CodeChunk}
\begin{CodeInput}
R>   devtools::install_github("zoometh/iconr", build_vignettes=TRUE)
\end{CodeInput}
\end{CodeChunk}
%

\subsection{External package} \label{sec:ext_pck}

The \pkg{decorr} package imports the following packages:
\begin{itemize}
\setlength\itemsep{.1em}
\item \pkg{magick} for image manipulation \citep{Ooms18}
\item \pkg{igraph} for graph and network analysis \citep{Csardi06}
\item \pkg{rgdal}  to read \emph{shapefiles} of nodes and edges \citep{Bivand19}
\item \pkg{grDevices} for colors and font plotting, \pkg{graphics} for graphics, \pkg{utils} and \pkg{methods} for formally defined methods and \emph{varia} methods (all combinations, etc.) \citep{R19}
\end{itemize}

\subsection{Data} \label{sec:ext_data}

A training dataset is stored in the \code{extdata} folder of the \pkg{decorr} package. It contains dataframes of nodes' coordinates, edges' coordinates, decorations' listing in different formats (\code{.tsv}, \code{.csv} or \emph{shapefiles}) and the decoration drawings (\code{.jpg}) of five Late Bronze Age stelae belonging to the so-called 'Warrior stelae' family -- with about 140 stelae -- mostly concentrated in the Southwest Iberian peninsula \citep{Celestino01}. \\

\smallbreak
\textbf{Imgs dataframe}
\\
The inventory of decorations is stored in the \code{imgs} dataframe (Tab. \ref{tab:timg}). The field \code{imgs$idf} is the short name of the decoration, useful during statistical analysis. The primary key of each decoration is the concatenate of \code{imgs$site} and \code{imgs$decor}. These keys will allow joints with the other dataframes (nodes and edges)
%
<<results='asis', echo=FALSE, warning=FALSE>>=
imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
                   sep="\t", stringsAsFactors = FALSE)
print(xtable::xtable(imgs,
                     caption = "The studied corpus, the \\code{imgs.tsv} dataframe",
                     label = "tab:timg",
                     size=6,
                     align=c(
                       "p{0.01cm}",
                       "c",
                       ">{\\raggedright\\arraybackslash\\rule{0pt}{1em}\\addtolength{\\baselineskip}{-1.5pt}}p{2.1cm}",
                       ">{\\raggedright\\arraybackslash\\addtolength{\\baselineskip}{-1.5pt}}p{2.1cm}",
                       "p{8.5cm}")),
      table.placement="!htbp",
      include.rownames=FALSE)
@
%

At first the drawings dataset \citep{DiazGuardamino10} can be checked by using the imgs dataframe and the \pkg{magick} package (Fig. \ref{fig:dataset})
%
<<timg, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=15, fig.height=11.2, fig.cap="\\label{fig:dataset}Decoration images of the training dataset">>=
pth <- system.file("extdata", package = "decorr")
imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
                   sep="\t", stringsAsFactors = FALSE)
lims <- list()
for(i in 1:nrow(imgs)){
  i1 <- image_read(paste0(pth,"\\",imgs[i,"img"]))
  lbl.txt <- paste0(imgs[i,"idf"], "\n",
                    imgs[i,"site"], "\n",
                    imgs[i,"decor"], "\n",
                    imgs[i,"img"], "\n",
                    image_info(i1)$width, "*", image_info(i1)$height, " px")
  i1 <- image_annotate(i1, lbl.txt, location = "northwest",
                       size = 28, color = "red")
  lims[[length(lims)+1]]<- i1
}
out.img <- image_append(c(image_append(c(lims[[1]], lims[[2]], lims[[3]])),
                          image_append(c(lims[[4]], lims[[5]]))),
                        stack = TRUE)
par(mar=c(0, 0, 0, 0))
plot(out.img)
@
%
To construct a graph overlapping the decoration images listed in the \code{images} dataframe, the first step is to load nodes and edges dataframes.
% \begin{itemize}
% \item The \code{nodes} dataframe
% \end{itemize}
\smallbreak
\textbf{Nodes dataframe}
\\
It contains the required minimum variables for the analysis (Tab. \ref{tab:tnds}).

<<tndss, results='asis', echo=TRUE>>=
caption <- "Nodes (from \\code{nodes.csv} dataframe)"
nodes <- read.table(system.file("extdata", "nodes.csv", package = "decorr"),
                    sep=";", stringsAsFactors = FALSE)
print(xtable::xtable(head(nodes),
                     caption = caption,
                     label = "tab:tnds"),
      include.rownames = FALSE,
      table.placement = "!htbp")
@

Since a prehistorical site can have various decorated objects, the primary key of the decoration is based on two fields: \code{nodes$site} and \code{nodes$decor}. The \code{nodes$id} is the identifier of the node. The \code{nodes$type} field is the default variable for further statistical analysis. Here, \code{nodes$type} refers to the typology of the nodes (anthropomorph, weapons, etc.). The \code{nodes$x} and \code{nodes$y} columns refer to the \emph{x} and \emph{}y coordinates of the nodes. As said, in the first place theses coordinates come from the GIS. But, in a GIS, the coordinates origin (0, 0) is the bottom-left corner and exist negative values, while this origin is top-left for any \proglang{R} matrices (rasters, grids, dataframes, etc.) with only positive values. To recover the local \emph{y} value of nodes and edges on the decoration image, the absolute value \emph{y} value and the image height, as a constant offset, are computed.
\smallbreak
% \begin{itemize}
% \item The \code{edges} dataframe
% \end{itemize}
\textbf{Edges dataframe}
\\
The edges dataframe is quite similar to the nodes dataframe (Table \ref{tab:teds}).

<<tedss, results='asis', echo=TRUE>>=
caption <- "Edges of Cerro Muriano 1 (from \\code{edges.csv} dataframe)"
edges <- read.table(system.file("extdata", "edges.csv", package = "decorr"),
                    sep=";",stringsAsFactors = FALSE)
edges.cm1 <- subset(edges, decor == "Cerro Muriano 1")
print(xtable::xtable(head(edges.cm1),
                     caption=caption,
                     label="tab:teds"),
      include.rownames = FALSE,
      table.placement="!htbp")
@

Fields \code{edges$site} and \code{edges$decor} are the primary key of decoration. The fields \code{edges$a} and \code{edges$b} are the equivalent to columns \emph{from} and \emph{to} in Graph theory. Even if undirected graphs will the most common in further studies, this direction helps to distinguish between nodes. The column \code{edges$a} is the identifier of \emph{starting node} or \emph{main node} or \emph{overlapping node}. The column \code{edges$b} is the identifier of the \emph{ending node} or \emph{attribue node} or \emph{overlapped node}. The column \code{edges$type} is the type of relation (normal, attribute, overlapping, etc.) between the node a and the node b. There is no need to record the coordinates of the edges, these coordinates are calculated from the nodes dataframe.

The \pkg{decorr} package functions can be divided into:
\begin{enumerate}
\item management functions
\item graphical functions
\item single decoration functions
\item comparisons functions
\end{enumerate}

Once these dataframes loaded, the list of decoration graphs can be calculated with \code{list_dec()}.

\subsection{Management functions} \label{sec:manage}

The management functions are \code{list\_dec()}, \code{contemp\_nds()} and \code{named\_elements()}

\begin{Code}
list_dec(imgs,
         nodes,
         edges)
\end{Code}

Once imgs, nodes and edges dataframes loaded, the \code{list\_dec()} function allows to create graphs for all decorations within these dataframes. The result is a list of \code{igraph} objects. As an example, the first graph of the list can be plotted with the \code{igraph::plot()} function (Fig. \ref{fig:igraph})

<<igraphs, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=2.5, fig.height=2.5, fig.cap="\\label{fig:igraph}R interface. Plot of the first graph of the list (Cerro Muriano 1 decoration)">>=
lgrph <- list_dec(imgs, nodes, edges)
par(mar=c(0, 0, 0, 0))
plot(lgrph[[1]],
     vertex.color = "orange",
     vertex.frame.color = "orange",
     vertex.label.color = "black",
     vertex.size = 14,
     vertex.label.cex = .8,
     edge.color = "orange",
     vertex.label.family = "Helvetica"
)
@

It occurs that some nodes are non-contemporaneous one with the others, like for the Ibahernando stelae. This stelae have been found reused as a funerary steale during Roman times \citep{Almagro66b}. A latin inscription (\code{ecriture}, node 1) have been carved over a spear (\code{lance}, node 2) and overlaps partially a V-notched shield (\code{bouclier}, node 3). The edges between node 1 and node 2, and the edge between node 1 and node 3, are \emph{overlap} edges (Figure \ref{fig:ibah}).

<<ibah, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=4, fig.asp=1, fig.cap="\\label{fig:ibah}Ibahernando decoration, node 1 overlaps node 2 and node 3">>=
par(mar=c(0,0,0,0))
nds.df <- read_nds(site = "Ibahernando", 
                   decor = "Ibahernando",
                   doss = system.file("extdata", 
                                      package = "decorr"))
eds.df <- read_eds(site = "Ibahernando",
                   decor = "Ibahernando",
                   doss = system.file("extdata", 
                                      package = "decorr"))
img.graph <- plot_dec_grph(nodes = nds.df,
                           edges = eds.df,
                           site = "Ibahernando",
                           decor = "Ibahernando",
                           lbl.size = 2,
                           doss = system.file("extdata", 
                                              package = "decorr"))
plot(magick::image_read(img.graph))
@

In this case, the non-contemporaneous layers of decoration should be removed before the comparison process. To do so, the original graph (1-component) will be splitted into different sub-graphs (\emph{n}-component) by selecting \code{>} edges. The studied graph component will be retrieved with the component membership of a selected node in the \code{contemp_nds()} function parameters.

To study only the Late Bronze Age iconographic layer of the Ibahernando steale, we can choose the Late Bronze Age node 4, the figure of sword (\code{epee}) dated to the middle and final stages of Late Bronze Age (ca 1250-950 BC). This node is supposed to be contemporaneous to node 2 and node 3 (Figure \ref{fig:ibah2}).

<<ibah2, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=8, fig.asp=1, fig.cap="\\label{fig:ibah2}Ibahernando stelae before and after the selection of node 4 (sword) graph component">>=
par(mar=c(0,0,0,0))
selected.nd <- 4
nds.df <- read_nds(site = "Ibahernando", 
                   decor = "Ibahernando",
                   doss = system.file("extdata", 
                                      package = "decorr"))
eds.df <- read_eds(site = "Ibahernando",
                   decor = "Ibahernando",
                   doss = system.file("extdata", 
                                      package = "decorr"))
l_dec_df <- contemp_nds(nds.df, eds.df, selected.nd)
Ibahernando <- plot_dec_grph(nodes = nds.df,
                             edges = eds.df,
                             site = "Ibahernando",
                             decor = "Ibahernando",
                             nd.var = "type",
                             lbl.color = "brown",
                             lbl.size=2.2,
                             doss = system.file("extdata", 
                                                package = "decorr"))
Ibahernando.img <- magick::image_read(Ibahernando)
Ibahernando.contemp <- plot_dec_grph(nodes = l_dec_df[[1]],
                                     edges = l_dec_df[[2]],
                                     site = "Ibahernando",
                                     decor = "Ibahernando",
                                     nd.var = "type",
                                     lbl.color = "brown",
                                     lbl.size=2.2,
                                     doss = system.file("extdata", 
                                                        package = "decorr"))
Ibahernando.contemp.img <- magick::image_read(Ibahernando.contemp)
out <- magick::image_append(c(Ibahernando.img, Ibahernando.contemp.img))
plot(out)
@

\begin{Code}
named_elements(grph,
               focus = "edges",
               nd.var = "type")
\end{Code}

The \code{named\_elements()} function allows to calculate the textual expression of graph elements (nodes and edges) with a disambiguation of these elements -- adding '#' suffixes -- when several nodes or edges have the same values.

\begin{CodeChunk}
\begin{CodeInput}
R> lgrph <- list_dec(imgs, nodes, edges)
R> as.character(named_elements(lgrph[[2]], focus = "edges", nd.var="type"))
## [1] "epee-=-fibule" "bouclier-=-miroir"
## [3] "epee-=-lance" "bouclier-=-chariot_char"
## [5] "chariot_char-+-cheval" "chariot_char-+-cheval#"
## [7] "arc-=-bouclier" "arc-=-chariot_char"
## [9] "arc-=-miroir" "bouclier-=-fibule"
## [11] "bouclier-=-epee" "epee-=-miroir"
## [13] "arc-+-fleche" "chariot_char-+-roue"
## [15] "chariot_char-+-roue#"
R> as.character(named_elements(lgrph[[2]], focus = "nodes", nd.var="type"))
## [1] "chariot_char" "bouclier" "arc" "miroir"
## [5] "fibule" "epee" "lance" "cheval"
## [9] "cheval#" "roue" "roue#" "fleche"
\end{CodeInput}
\end{CodeChunk}

\subsection{Graphical functions} \label{sec:functions_gr}

The \pkg{decorr} package has two graphical functions: \code{labels_shadow()} and \code{side_plot()}.

\begin{Code}
labels_shadow(x, y = NULL, labels,
              col = "black", bg = "white",
              theta = seq(0, 2 * pi, length.out = 50),
              r = 0.1,
              cex = 1, ...)
\end{Code}

\code{labels_shadow()} is a re-use of the \code{shadowtext()} function from the \pkg{TeachingDemos} package \citep{Snow20}. It plots labels (\code{labels}) at specific coordinates (\code{x, y}) with a contrasting buffer to make them more visible.

\begin{Code}
side_plot(grp, doss, nd.var, focus = "nodes",
          nd.color = c("orange", "red"),
          nd.size = c(0.5, 1),
          ed.color = c("orange", "red"),
          ed.width = c(1, 2),
          lbl.size = 0.5)
\end{Code}

\code{side_plot()} allows to plot two decoration graphs side-by-side for elements (nodes or edges) comparisons. It take a list of two graph (\code{grp}), the path of the working directory (\code{doss}) and the name of the studied node variable (\code{nd.var}) and return an image.

\subsection{Single decoration functions} \label{sec:functions_one}

Functions allowing to create a geometric graph for a single decoration are \code{read\_nds()} and \code{read\_eds()}. These functions allow to read respectively a file of nodes and a file of edges (\code{.csv}, \code{.tsv} or \code{.shp} files).

\begin{Code}
read_nds(site,
         decor,
         doss = getwd(),
         nodes = "nodes",
         dev = ".tsv")
\end{Code}

\code{read_nds()} is close to the \proglang{R} native \code{read.table()} function, but allows to read \emph{shapefiles} of nodes. For example, the nodes of the Torrejon Rubio 1 decoration can be read from the the \emph{shapefile}, and the 6\textsuperscript{th} node -- a sword (\code{epee}) -- displayed (Table \ref{tab:nd}).

<<nd, results='asis', echo=TRUE>>=
caption <- "Torrejon Rubio 1 decoration node 6: the sword."
sit <- "Torrejon Rubio" ; dec <- "Torrejon Rubio 1"
nds.df <- read_nds(site = sit, decor = dec, dev = ".shp",
                   doss = system.file("extdata", package = "decorr"))
print(xtable::xtable(nds.df[nds.df$id == 6, ],
                     caption = caption,
                     label = "tab:nd",
                     size = 7),
      include.rownames = FALSE,
      table.placement = "!htbp")
@

\begin{Code}
read_eds(site,
         decor,
         doss = getwd(),
         edges = "edges",
         nodes = "nodes",
         dev = ".tsv")
\end{Code}

\code{read_eds()} permits to read a \code{.csv} or \code{.tsv} dataframe, or a \emph{shapefiles}, of edges. When the data source is a dataframe, the coordinates of the edges are retrieved from the nodes dataframe (Table \ref{tab:ed}).

<<ed, results='asis', echo=TRUE >>=
caption <- paste0("First edge of the Torrejon Rubio 1 decoration, ",
                  "the proximity between the sword (node 6) ",
                  "and the fibula (node 5)")
eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
                   doss = system.file("extdata", package = "decorr"))
print(xtable::xtable(eds.df[1,],
                     caption = caption,
                     label = "tab:ed",
                     size = 7),
      include.rownames = FALSE,
      table.placement = "!htbp")
@

Once, the imgs, nodes and edges dataframes have been read, a geometric graph can be build and plotted over a specific decoration image with the \code{plot_dec_grph()} function.

\begin{Code}
plot_dec_grph(nodes = NULL,
              edges = NULL,
              site,
              decor,
              doss = getwd(),
              nd.var = 'id',
              nd.color = 'orange', nd.size = 1.7,
              lbl.color = 'black', lbl.size = 1.2,
              ed.color = c("orange", "blue"), ed.lwd = 4,
              img.format = "png")
\end{Code}

The parameters \code{nodes} and \code{edges} concern respectively the nodes and edges dataframes. The parameters \code{site} and \code{decor} allow to choose a specific decoration. The \code{nd.var} parameter allows to decide which field of the nodes will be displayed as the label, by default this is the \code{nodes$id} field. For example, we can plot the graph of the Torrejon Rubio 1 decoration with the type of the nodes (Figure \ref{fig:trst}).

<<trst, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=4, fig.asp=1, fig.cap="\\label{fig:trst}Torrejon Rubio 1 decoration">>=
par(mar=c(0, 0, 0, 0))
sit <- "Torrejon Rubio" ; dec <- "Torrejon Rubio 1"
nds.df <- read_nds(site = sit, decor = dec, dev = ".tsv",
                   doss = system.file("extdata", package = "decorr"))
eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
                   doss = system.file("extdata", package = "decorr"))
img.graph <- plot_dec_grph(nodes = nds.df,
                           edges = eds.df,
                           site = sit,
                           decor = dec,
                           doss = system.file("extdata", 
                                              package = "decorr"),
                           nd.var = "type", lbl.size = 2)
out <- magick::image_read(img.graph)
plot(out)
@

\subsection{Decoration comparison functions} \label{sec:functions_one}

These functions allow to compare different graph decorations. The function \code{list_compar()} is used to compare common nodes and common edges depending on the node variable \code{nd.var}.

\begin{Code}
list_compar(lgrph,
            nd.var = "type",
            verbose = FALSE)
\end{Code}

Whereas the \code{nd.var} (by default, the column \code{type}) refers to the nodes and can be any other column create by the user (e.g., \code{technique}, \code{color}, etc.), the comparisons always take into account the field \code{type} (\code{=}, \code{+}, \code{>}, ...) of the edges. Comparisons are done between one or more pairwise of decorations. In the training dataset, there are five (5) decorations in the default dataset, so there is $\frac{5!}{(5-2)!2!}={10}$ possible pairwise comparisons. This function calls \code{nds_compar()} and \code{eds_compar()}.

\begin{Code}
nds_compar(grphs,
           nd.var = "type")

eds_compar(grphs,
           nd.var = "type")
\end{Code}

The functions \code{nds_compar()} and \code{eds_compar()} take a pair of graphs \code{grphs} and return respectively the common nodes and the common edges based on the \code{nd.var} field of nodes (by default, the column \code{type}).

\smallbreak
The function \code{plot_compar()} allows to plot and save two figures side-by-side for one pairwise, or more, of decorations identifying common elements.

\begin{Code}
plot_compar(listg, graph2 = NULL, focus = "nodes",
            doss = getwd(),
            nd.color = c("orange", "red"), nd.size = c(0.5, 1),
            ed.color = c("orange", "red"), ed.width = c(1, 2),
            lbl.size = 0.5, img.format = "png", res = 300)
\end{Code}

\code{plot_compar()} takes a list of graphs \code{listg}, created previously by \code{list_compar()}, and a vector with the ids of the graphs to compare (\code{graph2}). The function creates an image for each pairwise comparison with the decorations plotted side-by-side with elements: common nodes (\code{focus = "nodes"}, by default) or common edges (\code{focus = "edges"}). For each pairwise, if an element exists only in one decoration, it is displayed in orange by default (\code{nd.color[1]} or \code{ed.color[1]}). When an element is present on the two decorations, it is displayed in red by default (\code{nd.color[2]} or \code{ed.color[2]}). The function returns the paths to the created images.

Let us compare the decoration 1 and the decoration 4 and identify common edges (Figure \ref{fig:ceds}).

<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=8, fig.height=4.2, fig.cap="\\label{fig:ceds}Comparisons between decoration 1 (Cerro Muriano 1) and decoration 4 (Zarza de Montsanchez)">>=
par(mar=c(0,0,0,0))
g.compar <- list_compar(lgrph, "type")
eds_compar <- plot_compar(g.compar, c(1, 4),
                          focus = "edges",
                          doss = system.file("extdata", package = "decorr"))
plot(magick::image_trim(magick::image_read(eds_compar)))
@

The comparison shows that the decoration 1 (Cerro Muriano 1) and the decoration 4 (Zarza de Montsanchez) have two (2) common edges: \code{lance -=- personnage} and \code{bouclier -=- personnage}.

The function \code{same_elements()} allows to count matching nodes and matching edges between decoration pairwises.

\begin{Code}
same_elements(lgrph,
              nd.var = "type",
              focus = "nodes")
\end{Code}

The \code{same_elements()} function take the list of graphs (\code{lgrph}) produced by \code{list_dec()} and the node variable on which the comparison will be done, by default \code{type} (\code{nd.var = "type"}). The result is a square matrix between all pairwise comparisons with the number of common elements (nodes or edges in the cells). The diagonal of the square matrix shows the total number of element for a given decoration. For example, we can compute the matrix of common edges (Table \ref{tab:mat.ed}).

<<results='asis', echo=TRUE>>=
df.same_edges <- same_elements(lgrph, "type", "edges")
caption <- "Number of common edges between all graph decoration pairwises"
print(xtable::xtable(df.same_edges,
                     caption = caption,
                     label = "tab:mat.ed",
                     size = 8,
                     digits = c(0)),
      table.placement="!htbp",
      include.rownames = TRUE)
@

The edges comparisons between the decoration 1 and the decoration 4 show that they have two (2) common edges, see also the Figure \ref{fig:ceds}.

\section{Illustrations} \label{sec:illustrations}

In order to demonstrate the main insight of a graph-based analysis of the decorations, we will compare two classifications. The first classification is based on the counts of common nodes. Since the precise location of the nodes is usualy not registred, his is the most commonly analysis applied on large series. The second classification is based on the presence of common edges.

At first, the Late Bronze Age phase of Ibahernando stelae must be selected (ie, the Roman latin writing must be removed). After the selection of the graph elements of Ibahernando stelae (decoration 5), nodes and edges of the node 2 connected component, the lance (spear), is selected with \code{contemp_nds()}, a new graph object is created (\code{grph}) and used to update the list of graphs (\code{lgrph}).

<<results='asis', echo=TRUE, fig.height= 5>>=
dataDir <- system.file("extdata", package = "decorr")
nds.df <- read_nds(site = "Ibahernando",
                   decor = "Ibahernando",
                   doss = dataDir)
eds.df <- read_eds(site = "Ibahernando",
                   decor = "Ibahernando",
                   doss = dataDir)
l_dec_df <- contemp_nds(nds.df, eds.df, selected.nd = 2)
l_dec_df[[1]]$site <- l_dec_df[[1]]$decor <- NULL
l_dec_df[[2]]$site <- l_dec_df[[2]]$decor <- NULL
grph <- graph_from_data_frame(l_dec_df[[2]], vertices = l_dec_df[[1]],
                           directed = FALSE)
grph$site <- grph$decor <- "Ibahernando"
grph$name <- 5
lgrph[[5]] <- grph
@

Then, \code{same_elements()} is done on nodes and edges (Table \ref{tab:common.elts})

<<results='asis', echo=TRUE>>=
df.same_edges <- same_elements(lgrph, "type", "edges")
df.same_nodes<- same_elements(lgrph, "type", "nodes")
@

\begin{table}[!htbp]
% \begin{table}[H]
\centering
\begin{minipage}{.3\textwidth}
\centering
\textbf{Common nodes}\raisebox{-1ex}{}\\
<<echo=FALSE,results='asis' >>=
print(xtable::xtable(df.same_nodes,
                     size = 8,
                     digits = c(0)),
      floating=FALSE,
      include.rownames = TRUE)
@
\end{minipage}
\begin{minipage}{.3\textwidth}
\centering
\textbf{Common edges}\raisebox{-1ex}{}\\
<<echo=FALSE,results='asis'>>=
print(xtable::xtable(df.same_edges,
                     size = 8,
                     digits = c(0)),
      floating=FALSE,
      include.rownames = TRUE)
@
\end{minipage}
\caption{Tables showing the number of common nodes and common edges between each pair of decorations}
\label{tab:common.elts}
\end{table}

% the inverse values of common nodes and common edges are calculated with the function \code{inv()} from the package \pkg{matlib} \citep{Friendly20}, and then 
Once the heatmap matrices calculated, the distance matrices are calculated with the native \code{dist()} function \citep{R19} (Table \ref{tab:dist.elts}).

\begin{table}[!htbp]
% \begin{table}[H]
\centering
\begin{minipage}{.4\textwidth}
\centering
\textbf{Distance matrix on nodes}\raisebox{-1ex}{}\\
<<echo=FALSE,results='asis', warning=F >>=
library("matlib")
dist.nodes  <- dist(as.matrix(df.same_nodes))
mat.nodes <- as.matrix(dist.nodes)
mat.nodes[lower.tri(mat.nodes, diag = TRUE)] <- NA
print(xtable::xtable(mat.nodes,
                     size = 8,
                     digits = c(2)),
      floating=FALSE,
      include.rownames = TRUE)
@
\end{minipage}
\begin{minipage}{.4\textwidth}
\centering
\textbf{Distance matrix on edges}\raisebox{-1ex}{}\\
<<echo=FALSE,results='asis'>>=
dist.edges <- dist(as.matrix(df.same_edges))
mat.edges <- as.matrix(dist.edges)
mat.edges[lower.tri(mat.edges, diag = TRUE)] <- NA
print(xtable::xtable(mat.edges,
                     size = 8,
                     digits = c(2)),
      floating=FALSE,
      include.rownames = TRUE)
@
\end{minipage}
\caption{Distance matrices on nodes and common edges between each pair of decorations}
\label{tab:dist.elts}
\end{table}

For both nodes and edges, the results show that the most important differences are between the decoration 1 and the decoration 2: for nodes, dist = \Sexpr{round(max(mat.nodes, na.rm = T), 2)} ; for edges, dist =  \Sexpr{round(max(mat.edges, na.rm = T), 2)}. At the opposite, the decorations showing the more similar compositions are decoration 3 and decoration 5: for nodes, dist = \Sexpr{round(min(mat.nodes, na.rm = T), 2)} ; for edges, dist =  \Sexpr{round(min(mat.edges, na.rm = T), 2)}. The compared decorations, with their common elements, can be plotted with \code{plot_compar()}.
A hierachical clustering (HC) of the distance matrices, with the native function \code{hclust()} \citep{R19}, summarize these similarites/dissimilarities (Figure \ref{fig:hclust})

<<hclust, echo=TRUE, warning=FALSE, fig.height=5, fig.pos = '!htbp', fig.cap="\\label{fig:hclust}Hierachical clusterings on decoration common elements">>=
par(mfrow=c(1, 2))
hc1 <- hclust(dist.nodes, method = "ward.D")
hc2 <- hclust(dist.edges, method = "ward.D") 
plot(hc1, main = "common nodes")
plot(hc2, main = "common edges")
@

Even if in this example the clustering of decorations on common nodes and the clustering of decorations on common edges give the same results, another distribution of the same nodes, could give two different clusterings. The clustering on common nodes and the clustering on common edges can be compared with the \code{tanglegram()} function from the \pkg{dendextend} package \citep{Galili15} (Figure \ref{fig:hclust2})

<<hclust2, echo=TRUE, warning=FALSE, fig.height=3, fig.pos = '!htbp', fig.cap="\\label{fig:hclust2}Comparison of hierachical clusterings on decoration common elements, nodes (left) and edges (right)">>=
suppressPackageStartupMessages(library(dendextend))
suppressPackageStartupMessages(library(dplyr))
dend1 <- as.dendrogram (hc1)
dend2 <- as.dendrogram (hc2)
dendlist(dend1, dend2) %>%
  untangle(method = "step1side") %>% 
  tanglegram()                       
@

%% -- Summary/conclusions/discussion -------------------------------------------

\section{Summary and discussion} \label{sec:summary}

The \pkg{decorr} framework is grounded on a parsimonious model, where Tobler's \emph{near things} \citep{Tobler70} of any iconographical content is recorded with a minimal of \emph{a priori} definitions. Those definitions only concern the graph elements: graphical units variables (ie, nodes variables) and types of relations they share (edges). In this example we propose the iconographical \code{type} (\code{personnage}, \code{casque}, etc.) nodes as the studied variable (\code{nd.var}), but the user of the package can create and choose any other study variable: \code{technique}, \code{color}, etc. The current edge notation with \emph{normal} edge, \emph{attribute} edge and \emph{overlapping} edge allow to describe most cases of Prehistorical iconography, but can also be extended to accept new types like, for example, the \emph{interact} edge when two graphical units interact (ie, when an action is visible). Currently, the undecorated parts of the support -- the background -- is inactive and considered as homogeneous it is possible to consider but a crack on a rock or a carene on a pottery \citep{Huet18a}, like relevant topologic entities to understand the iconographical composition. Then a new class of nodes, beside the graphical units, can be registred to create a multipartite graph (\emph{n}-mode graph). The plasticity of graph theory will allow to filter these different classes of nodes.

Beside giving a parsimonious framework to register, manage and analyse compositions, for further statistical analysis, the \pkg{decorr} package allows also to avoid, empirical and idiosynchratic top-down hierachical typologies (e.g., a category of graphical units with \emph{n} types, a type with \emph{n} subtypes, a subtype with \emph{n} varieties, etc.) that lead also to tedious inventories and difficult possible comparisons \emph{within} and \emph{between} these categories. Indeed, to quote the GUs it is more parsimonious to use the edge notation rather than multiplying hierarchical categories. For example, with the two types of GUs \code{casque} (helmet) and \code{personnage} (anthropomorphe) and the \emph{attribute} edge it is possible to describe: (1) a case where an anthropomorph wears a helmet (like for the Cerro Muriano steale, Figure \ref{fig:dec.1.4} left), (2) a case where an anthropomorph is separated from a helmet (like for the Zarza de Montanchez, Figure \ref{fig:dec.1.4} right). With a top-down hierachical typology, three types of GUs would have been required: (1) anthropomorph wearing a helmet (Cerro Muriano); (2) anthropomorph and (3) helmet (Zarza de Montanchez).

<<dec.1.4, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=15, fig.asp=1, fig.cap="\\label{fig:dec.1.4}Ibahernando stelae before and after the selection of node 4 (sword) graph component">>=
par(mar=c(0, 0, 0, 0))
nds.df <- read_nds(site = "Cerro Muriano", 
                   decor = "Cerro Muriano 1",
                   doss = system.file("extdata", 
                                      package = "decorr"))
eds.df <- read_eds(site = "Cerro Muriano",
                   decor = "Cerro Muriano 1",
                   doss = system.file("extdata", 
                                      package = "decorr"))
cm1 <- plot_dec_grph(nodes = nds.df,
                     edges = eds.df,
                     site = "Cerro Muriano",
                     decor = "Cerro Muriano 1",
                     nd.var = "type",
                     lbl.size = 1.7,
                     doss = system.file("extdata", 
                                        package = "decorr"))
cm1.img <- magick::image_read(cm1)
nds.df <- read_nds(site = "Zarza de Montanchez", 
                   decor = "Zarza De Montanchez",
                   doss = system.file("extdata", 
                                      package = "decorr"))
eds.df <- read_eds(site = "Zarza de Montanchez",
                   decor = "Zarza De Montanchez",
                   doss = system.file("extdata", 
                                      package = "decorr"))
zm <- plot_dec_grph(nodes = nds.df,
                            edges = eds.df,
                            site = "Zarza de Montanchez",
                            decor = "Zarza De Montanchez",
                            nd.var = "type",
                            lbl.size = 1.7,
                            doss = system.file("extdata", 
                                               package = "decorr"))
zm.img <- magick::image_read(zm)
out.img <- magick::image_append(c(cm1.img, zm.img))
plot(out.img)
@

Archaeologists use quite commonly the Harris matrix to model the relative chronology of archaeological layers (in a case of an excavation) or the relative chronology of iconographical layers (in case of a decorated support). The \emph{overlap} edge, by recording the anterior-posterior relation between pairwises of graphical units, is first operator permiting to trace the Harris matrix directly of the studied decoration.

In Archaeology, the iconography has been considered as a privileged elements to the identification of cultural facies \citep{Renfrew91}. By integrating the recording, the management and the analysis process in a simple manner, within the framework of graph Theory and binary topological telationships (\emph{birel}), the \pkg{decorr} framework opens possibilities of studying graphical systems with normalized indexes over large series (Later Paleolithic cave paintings, \emph{cups-and-rings} art, Valcamonica Iron Age \emph{pittoti}, etc.) a long period of time, and on heterogeneous source data.

\section*{Acknowledgments}

A part of the IT development of the \pkg{decorr} package has been done in the frame of a postdoctoral position (2015-16) supported by LabEx ARCHIMEDE from "Investissement d'Avenir" program ANR-11-LABX-0032-01.

%% -- Bibliography -------------------------------------------------------------
%% - References need to be provided in a .bib BibTeX database.
%% - All references should be made with \cite, \citet, \citep, \citealp etc.
%%   (and never hard-coded). See the FAQ for details.
%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
%% - Titles in the .bib should be in title case.
%% - DOIs should be included where available.

\bibliography{bibref}

% 
% %% -- Appendix (if any) --------------------------------------------------------
% %% - After the bibliography with page break.
% %% - With proper section titles and _not_ just "Appendix".
% 
% \newpage
% 
% \begin{appendix}
% 
% \section{More technical details} \label{app:technical}
% 
% \begin{leftbar}
% Appendices can be included after the bibliography (with a page break). Each
% section within the appendix should have a proper section title (rather than
% just \emph{Appendix}).
% 
% For more technical style details, please check out JSS's style FAQ at
% \url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
% which includes the following topics:
% \begin{itemize}
% \item Title vs.\ sentence case.
% \item Graphics formatting.
% \item Naming conventions.
% \item Turning JSS manuscripts into \proglang{R} package vignettes.
% \item Trouble shooting.
% \item Many other potentially helpful details\dots
% \end{itemize}
% \end{leftbar}
% 
% 
% \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}
% 
% \begin{leftbar}
% References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
% references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
% \verb|\citealp| etc.\ (and never hard-coded). This commands yield different
% formats of author-year citations and allow to include additional details (e.g.,
% pages, chapters, \dots) in brackets. In case you are not familiar with these
% commands see the JSS style FAQ for details.
% 
% Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
% when acquiring the entries automatically from mixed online sources. However,
% it is important that informations are complete and presented in a consistent
% style to avoid confusions. JSS requires the following format.
% \begin{itemize}
% \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
% be used in the references.
% \item Titles should be in title case.
% \item Journal titles should not be abbreviated and in title case.
% \item DOIs should be included where available.
% \item Software should be properly cited as well. For \proglang{R} packages
% \code{citation("pkgname")} typically provides a good starting point.
% \end{itemize}
% \end{leftbar}
% 
% \end{appendix}

%% -----------------------------------------------------------------------------

\end{document}